---
title: "Time to go"
output: html_notebook
---

When should a fan leave a baseball game based solely on the score? Ever since I was a child, my father and I have disagreed over this point. His philosophy is that it's best to leave a game as soon as it's obvious who is going to win in order to beat the traffic. We've left too many games only to miss a late comeback. This made me curious about when we really should be leaving.

```{r prep, echo = TRUE, message=F, error=FALSE}
library(tidyverse)
library(stringr)
library(rvest)
library(lubridate)

```

Baseball data come from Retrosheet, which requires the following attribution:
"The information used here was obtained free of charge from and is copyrighted by Retrosheet.  Interested parties may contact Retrosheet at [www.retrosheet.org](www.retrosheet.org)." For convenice, I took the column names from the source code of the `retrosheet` package.

```{r data, echo = TRUE}
### Acquire and clean data

# Unzip files to the working directory
get_retrosheet <- function() {
# List all hyperlinks on page
links <- read_html("http://www.retrosheet.org/gamelogs/") %>%
html_nodes("a") %>%
html_attr("href")

# Only keep links to aggregated game files
links <-
links[str_detect(
links,
"http://www.retrosheet.org/gamelogs/gl[0-9][0-9][0-9][0-9]_[0-9][0-9].*zip"
)]
links <- links[!is.na(links)]

# Download and unzip all files
lapply(links, function(x) {
file_name <- basename(x)
download.file(x, file_name)
unzip(file_name)
TRUE
})

file_list <- list.files(pattern = "GL[0-9][0-9][0-9][0-9]\\.TXT$")
dataset <-
lapply(file_list, function(x)
read_csv(x, col_names = FALSE)) %>%
do.call("rbind", .) %>%
tbl_df()

# Column names from the retrosheet package
# I use all of the names here even though I ultimately ignore most of the variables
names(dataset) <-
c(
"Date",
"DblHdr",
"Day",
"VisTm",
"VisTmLg",
"VisTmGNum",
"HmTm",
"HmTmLg",
"HmTmGNum",
"VisRuns",
"HmRuns",
"NumOuts",
"DayNight",
"Completion",
"Forfeit",
"Protest",
"ParkID",
"Attendance",
"Duration",
"VisLine",
"HmLine",
"VisAB",
"VisH",
"VisD",
"VisT",
"VisHR",
"VisRBI",
"VisSH",
"VisSF",
"VisHBP",
"VisBB",
"VisIBB",
"VisK",
"VisSB",
"VisCS",
"VisGDP",
"VisCI",
"VisLOB",
"VisPs",
"VisER",
"VisTER",
"VisWP",
"VisBalks",
"VisPO",
"VisA",
"VisE",
"VisPassed",
"VisDB",
"VisTP",
"HmAB",
"HmH",
"HmD",
"HmT",
"HmHR",
"HmRBI",
"HmSH",
"HmSF",
"HmHBP",
"HmBB",
"HmIBB",
"HmK",
"HmSB",
"HmCS",
"HmGDP",
"HmCI",
"HmLOB",
"HmPs",
"HmER",
"HmTER",
"HmWP",
"HmBalks",
"HmPO",
"HmA",
"HmE",
"HmPass",
"HmDB",
"HmTP",
"UmpHID",
"UmpHNm",
"Ump1BID",
"Ump1BNm",
"Ump2BID",
"Ump2BNm",
"Ump3BID",
"Ump3BNm",
"UmpLFID",
"UmpLFNm",
"UmpRFID",
"UmpRFNm",
"VisMgrID",
"VisMgrNm",
"HmMgrID",
"HmMgrNm",
"WinPID",
"WinPNm",
"PID",
"PNAme",
"SavePID",
"SavePNm",
"GWinRBIID",
"GWinRBINm",
"VisStPchID",
"VisStPchNm",
"HmStPchID",
"HmStPchNm",
"VisBat1ID",
"VisBat1Nm",
"VisBat1Pos",
"VisBat2ID",
"VisBat2Nm",
"VisBat2Pos",
"VisBat3ID",
"VisBat3Nm",
"VisBat3Pos",
"VisBat4ID",
"VisBat4Nm",
"VisBat4Pos",
"VisBat5ID",
"VisBat5Nm",
"VisBat5Pos",
"VisBat6ID",
"VisBat6Nm",
"VisBat6Pos",
"VisBat7ID",
"VisBat7Nm",
"VisBat7Pos",
"VisBat8ID",
"VisBat8Nm",
"VisBat8Pos",
"VisBat9ID",
"VisBat9Nm",
"VisBat9Pos",
"HmBat1ID",
"HmBat1Nm",
"HmBat1Pos",
"HmBat2ID",
"HmBat2Nm",
"HmBat2Pos",
"HmBat3ID",
"HmBat3Nm",
"HmBat3Pos",
"HmBat4ID",
"HmBat4Nm",
"HmBat4Pos",
"HmBat5ID",
"HmBat5Nm",
"HmBat5Pos",
"HmBat6ID",
"HmBat6Nm",
"HmBat6Pos",
"HmBat7ID",
"HmBat7Nm",
"HmBat7Pos",
"HmBat8ID",
"HmBat8Nm",
"HmBat8Pos",
"HmBat9ID",
"HmBat9Nm",
"HmBat9Pos",
"Additional",
"Acquisition"
)

# Retrosheet uses "x" when the bottom of the 9th was not played
# because the home team was winning. Replace those "x" values
# with 0.
dataset$HmLine <- str_replace_all(dataset$HmLine, "x", 0)
dataset$VisLine <- str_replace_all(dataset$VisLine, "x", 0)

# Parse the box score
# Each digit is equal to the number of runs scored in one inning.
# The problem is that some innings see more than 9 runs scored.
# This could be coded in many ways that would be much faster,
# but for now it's a for loop that replaces all (xx) values with
# single letters. That enables the use of strsplit to attribute
# runs to the appropriate innings. The values are changed back
# to their numerical equivilents later in the code.
for (i in 10:47) {
dataset$HmLine <-
str_replace_all(dataset$HmLine,
paste0("\\(", i, "\\)"),
paste0("\\", rawToChar(as.raw(i))))
dataset$VisLine <-
str_replace_all(dataset$VisLine,
paste0("\\(", i, "\\)"),
paste0("\\", rawToChar(as.raw(i))))
}

# Create the variable gamecode as the unique combination of four variables
dataset <- dataset %>%
mutate(gamecode = paste(Date, VisTm, HmTm, DblHdr, sep = "_")) %>%
select(gamecode, Date, VisRuns, HmRuns, HmLine, VisLine) %>%
mutate(Date = ymd(Date)) %>%
mutate(winner = "tie") %>%
filter(!is.na(HmLine)) %>%
filter(!is.na(VisLine)) %>%
arrange(desc(Date))

# Remove extra innings.
# You're not going to leave a game in overtime, are you?
dataset$HmLine <- str_sub(dataset$HmLine, start = 1L, end = 9L)
dataset$VisLine <- str_sub(dataset$VisLine, start = 1L, end = 9L)

# Create one column for each half inning.
# inning1-1 is the top of the first,
# inning3-2 is the bottom of the third, etc. 
hmline <-
str_split(dataset$HmLine, pattern = "", simplify = TRUE) %>% tbl_df()
names(hmline) <- paste0("inning", 1:9, "-2")
visline <-
str_split(dataset$VisLine, pattern = "", simplify = TRUE) %>% tbl_df()
names(visline) <- paste0("inning", 1:9, "-1")

for (i in 10:47) {
hmline <-
apply(hmline, MARGIN = 2, function(x)
str_replace_all(x, paste0("\\", rawToChar(as.raw(
i
))), i))
visline <-
apply(visline, MARGIN = 2, function(x)
str_replace_all(x, paste0("\\", rawToChar(as.raw(
i
))), i))
}

# Replace the original box scores with the formatted scores
dataset <- select(dataset,-HmLine,-VisLine)
dataset <- cbind(dataset, hmline, visline)

# Convert all runs to numeric
dataset[, 6:23] <- apply(dataset[, 6:23], MARGIN = 2, as.numeric)

# How many runs did each team score without counting extra innings?
# In hindsight I could have used rowsum here
dataset$VisRunsReg <-
apply(dataset[, 15:23],
MARGIN = 1,
FUN = sum,
na.rm = TRUE)

dataset$HmRunsReg <-
apply(dataset[, 6:14],
MARGIN = 1,
FUN = sum,
na.rm = TRUE)
dataset <- tbl_df(dataset)

# Assign the winner column appropriately
dataset$winner[dataset$HmRuns > dataset$VisRunsReg] <- "home"
dataset$winner[dataset$HmRuns < dataset$VisRunsReg] <- "vis"

# Create columns for each half-inning showing how many runs
# the winner and loser scored in each half-inning.
# By definition at least half of these columns will be 0.
dataset[26:43] <- 0
names(dataset)[26:34] <- paste0("inning", 1:9, "-win")
dataset[dataset$winner == "home", 26:34] <-
dataset[dataset$winner == "home", 6:14]
dataset[dataset$winner == "vis", 26:34] <-
dataset[dataset$winner == "vis", 15:23]

names(dataset)[35:43] <- paste0("inning", 1:9, "-lose")
dataset[dataset$winner == "vis", 35:43] <-
dataset[dataset$winner == "vis", 6:14]
dataset[dataset$winner == "home", 35:43] <-
dataset[dataset$winner == "home", 15:23]

# Drop most columns and convert from wide to long
dataset <- dataset %>%
  select(1, 5, 26:43) %>%
  reshape2::melt(id.vars = c("gamecode", "winner")) %>%
  mutate(variable = ordered(variable)) %>% # Make innings an ordered factor
  arrange(gamecode, variable) %>% # Important for downstream analysis
  separate(variable, into = c("inning", "team"), sep = "-") %>%
  mutate(inning = str_sub(inning, -1, -1)) %>%
  mutate(inning = as.numeric(inning)) %>%
  tbl_df()
  
  
dataset$team <- "tie"
dataset$team[grepl("win", dataset$variable)] <- "win"
dataset$team[grepl("lose", dataset$variable)] <- "lose"

return(dataset)
}

games <- get_retrosheet()

```

Now that we have the data, let's look for trends that might affect our analysis. For example, how common are runs in a baseball game? Has scoring changed since the 19th century?

```{r eda}
### Exploratory data analysis

```

It's time to create an algorithm to determine the probability of a reversal when given a properly formatted dataset of scoring events throughout a collection of games. I'm naming this function "dad" because my father has traditionally been the one to determine when it's time to leave.

```{r dad}
dad <- function(dataset) {
  
  if(!is.data.frame(dataset)) {
    stop("dataset must be dataframe or tibble")
  }
  if(length(unique(dataset$gamecode)) < 2) {
    stop("dataset must contain data from multiple games")
  }
  
  dataset <- dataset %>%
    group_by(gamecode, team) %>%
    mutate(score = cumsum(value)) %>%
    filter(winner != "tie") %>%
    group_by(gamecode, inning) %>%
    arrange(team) %>%
    summarise(win_lead = first(value, order_by = team))
    
  dataset %>%
    group_by(gamecode, team) %>%
    mutate(score = cumsum(value)) %>%
    filter(winner != "tie") %>%
    filter(inning == 9)

}
```


```{r analysis}
### Perform analysis

```

## Future development
There are numerous potential future areas for the development of this project. The most intreguing is the addition of other sports.

### Hockey

### NASCAR

### Basketball
BasketballValue.com has event-level NBA data available for download covering the 2005-2006 through 2011-2012 seasons: http://basketballvalue.com/downloads.php

Basketball Reference has scoring-level data, but it's not in a format that's machine-readable: http://www.basketball-reference.com/play-index/plus/event_finder.cgi?request=1&event_code=fg&year_id=2017&is_playoffs=N&team_id=PHI

Drew Dresser has created a Python script for downloading ESPN basketball data: https://gist.github.com/drewdresser/8153408

### Soccer
OpenFootball has a massive database of soccer data. Its website references event-level data, but I haven't yet had a chance to explore: http://openfootball.github.io/news/

Here's a broader view of available soccer data sources: https://www.jokecamp.com/blog/guide-to-football-and-soccer-data-and-apis/

### Elections
Maybe this approach could apply to presidential or Congressional elections. Given the relatively small number of elections that have accurate tracking data, I'd only include election data for entertainment value.